### 2023-06-06 proofs/tex/anti-unification-lemmas.tex
anti-unification: & → ∪

\begin{lemma}[Soundness of Anti-Unification] \label{lemma:anti-unification-soundness}
    \hfill
    \begin{itemize}
        \item [$+$]  Assuming $[[iP1]]$ and $[[iP2]]$ are normalized,
        if $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
        then 
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uQ]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$, and
            \item $[[ [ausi] uQ = iPi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}

        \item [$-$] Assuming $[[iN1]]$ and $[[iN2]]$ are normalized,
        if $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
        then
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uM]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$, and
            \item $[[ [ausi] uM = iNi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove it by induction on 
    $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
    and mutually, $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$.
    Let us consider the last rule applied to infer this judgement.
    \begin{caseof}
        \item \ruleref{\ottdruleAUNVarLabel}, then $[[iN1]] = [[α⁻]] = [[iN2]]$,
              $[[Ξ]] = [[·]]$, $[[uM]] = [[α⁻]]$, and $[[aus1]] = [[aus2]] = [[·]]$.
            \begin{enumerate}
                \item $[[Γ ; · ⊢ α⁻]]$ follows from the assumption $[[Γ ⊢ α⁻]]$,
                \item $[[Γ ; · ⊢ · : ·]]$ holds trivially, and
                \item $[[ [·] α⁻ = α⁻ ]]$ holds trivially.
            \end{enumerate}
        \item \label{case:anti-unification-soundness:shiftu}
         \ruleref{\ottdruleAUShiftULabel}, then $[[iN1]] = [[↑iP1]]$,
                $[[iN2]] = [[↑iP2]]$, and the algorithm makes the recursive call:
                $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
                returning $[[(Ξ, ↑uQ, aus1, aus2)]]$ as the result.

@@

                Notice, that the induction hypothesis is applicable to 
                $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$:
                $[[iP1]]$ and $[[iP2]]$ are normalized, since $[[iN1]] = [[iP1 → iN'1]]$
                and $[[iN2]] = [[iP2 → iN'2]]$ are normalized.
                Similarly, the induction hypothesis is applicable to
                $[[Γ ⊨ iN'1 ≈au iN'2 ⫤ (Ξ', uM, aus'1, aus'2)]]$.

@@

                    \item $[[Γ ; · ⊢ ausi : Ξ]]$ and $[[Γ ; · ⊢ aus'i : Ξ']]$, 
                           and hence, $[[Γ ; · ⊢ ausi ∪ aus'i : Ξ ∪ Ξ']]$ (\cref{todo});

                    \item $[[ [ausi] uQ = iPi ]]$ and $[[ [aus'i] uM = iN'i ]]$.
                    Since $[[ausi ∪ aus'i]]$ restricted to $[[Ξ]]$ is $[[ausi]]$,
                    we have $[[ [ausi ∪ aus'i] uQ = iPi ]]$ and 
                    $[[ [ausi ∪ aus'i] uM = iN'i ]]$, and thus, 
                    $[[ [ausi ∪ aus'i] uQ → uM = iP1 → iN'1 ]]$
                \end{enumerate}

@@

        \item \ruleref{ottdruleAUArrowLabel}
        In this case, $[[Ξ]] = [[Ξ' ∪ Ξ'']]$, $[[aus1]] = [[aus'1 & aus''1]]$, and 
        $[[aus2]] = [[aus'2 & aus''2]]$,
        where the property holds for ($[[Ξ']]$, $[[aus'1]]$, $[[aus'2]]$) and 
        ($[[Ξ'']]$, $[[aus''1]]$, $[[aus''2]]$) by the induction hypothesis.
        Then since the merge of solutions does not change the types the variables are mapped to,
        the required property holds for $[[Ξ]]$, $[[aus1]]$, and $[[aus2]]$.

@@

\begin{lemma}[Completeness of Anti-Unification] \label{lemma:anti-unification-completeness}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            there exists $[[(Ξ', uQ', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$, and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

@@

        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            there exists $[[(Ξ', uM', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

@@

            Moreover, $[[ [aus'i]uQ' = iPi ]]$ and $[[ [aus'i]uM'' = iN''i ]]$, which means that
            $[[(Ξ', uQ', aus'1, aus'2)]]$ is an anti-unifier of $[[iP1]]$ and $[[iP2]]$,
            and $[[(Ξ', uM'', aus'1, aus'2)]]$ is an anti-unifier of $[[iN''1]]$ and $[[iN''2]]$.
            Then by the induction hypothesis, 
            $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ1, uQ, aus1, aus2)]]$ and 
            $[[Γ ⊨ iN''1 ≈au iN''2 ⫤ (Ξ2, uM, aus3, aus4)]]$ succeed.
            The result of the algorithm is $[[(Ξ1 ∪ Ξ2, uQ → uM, aus1 & aus3, aus2 & aus4)]]$,
            and it is left to show that $[[aus1 & aus3]]$ and $[[aus2 & aus4]]$ are defined.

            From \cref{obs:names-defined-by-mapping}, we know that 
            if $[[β̂⁻ ∊ dom(aus1) ∩ dom(aus3)]]$ then $[[ [aus1]β̂⁻ = [aus3]β̂⁻ ]]$,
            and if $[[β̂⁻ ∊ dom(aus2) ∩ dom(aus4)]]$ then $[[ [aus2]β̂⁻ = [aus4]β̂⁻ ]]$.
            This way, $[[aus1 & aus3]]$ and $[[aus2 & aus4]]$ are defined.

@@

\begin{lemma}[Initiality of Anti-Unification] \label{lemma:anti-unification-initial}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
            then $[[(Ξ, uQ, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uQ', aus'1, aus'2)]]$, i.e.
            if 
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$, and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uQ')]]$ and $[[ [ausr] uQ' = uQ ]]$. 
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$, 
            then $[[(Ξ, uM, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uM', aus'1, aus'2)]]$, i.e.
            if
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ and $[[ [ausr] uM' = uM ]]$.
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    First, let us assume that $[[uM']]$ is a metavariable $[[α̂⁻]]$. 
    Then we can take $[[ausr]] = [[α̂⁻]] \mapsto [[uM]]$, which satisfies the required properties:
    \begin{itemize}
        \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds since 
        $[[Ξ' | uv uM']] = [[{α̂⁻}]]$ and $[[Γ ; Ξ ⊢ uM]]$ by the soundness of anti-unification (\cref{lemma:anti-unification-soundness});
        \item $[[ [ausr] uM' = uM ]]$ holds by construction
        \item $[[ [ausr]α̂⁻]] = [[uM]]$ is the anti-unifier of 
            $[[iN1]] = [[ [aus'1] α̂⁻]]$ and $[[iN2]] = [[ [aus'2] α̂⁻]]$
            in context $[[Γ]]$, and hence, it is uniquely determined by them (\cref{obs:anti-unification-deterministic}).
    \end{itemize}

@@

        Then by the completeness of anti-unification (\cref{lemma:anti-unification-completeness}),
        the anti-unification algorithm succeeds on $[[iP1]]$ and $[[iP2]]$:
        $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ1, uQ, aus1, aus2)]]$;
        and on $[[iN'1]]$ and $[[iN'2]]$:
        $[[Γ ⊨ iN''1 ≈au iN''2 ⫤ (Ξ2, uM''', aus3, aus4)]]$.
        Notice that $[[aus1 & aus3]]$ and $[[aus2 & aus4]]$ are defined, 
        in other words, for any $[[β̂⁻]] \in [[Ξ1]] \cap [[Ξ2]]$,
        $[[ [aus1] β̂⁻ = [aus2] β̂⁻ ]]$ and $[[ [aus3] β̂⁻ = [aus4] β̂⁻ ]]$,
        which follows immediately from \cref{obs:names-defined-by-mapping}.
        This way, the algorithm proceeds by applying \ruleref{\ottdruleAUArrowLabel} and returns
        $[[(Ξ1 ∪ Ξ2, uQ → uM''', aus1 & aus3, aus2 & aus4)]]$.

@@

        Let us show that $[[ausr]] = [[ausr1 & ausr2]]$ is defined, 
        in other words, for any $[[β̂⁻]] \in [[(Ξ' | uv uQ')]] \cap [[(Ξ' | uv uM'')]]$,
        $[[ [ausr1] β̂⁻ = [ausr2] β̂⁻ ]]$.It holds because by the induction hypothesis,
        $[[ [ausri] β̂⁻ ]]$ is uniquely determined by $[[ [aus'1] β̂⁻ ]]$, $[[ [aus'2] β̂⁻ ]]$, and $[[Γ]]$,
        none of which depends on $i$.

        Let us show that $[[ausr]] = [[ausr1 & ausr2]]$ satisfies the required properties:
        \begin{itemize}
            \item $[[Γ ; Ξ1 ∪ Ξ2 ⊢ ausr1 & ausr2 : (Ξ' | uv uM')]]$ holds since 
            $[[Ξ' | uv uM']] = [[Ξ' | uv uQ' → uM'']] = [[(Ξ' | uv uQ') ∪ (Ξ' | uv uM'')]]$,
            $[[Γ ; Ξ1 ⊢ ausr1 : (Ξ' | uv uQ')]]$ and $[[Γ ; Ξ2 ⊢ ausr2 : (Ξ' | uv uM'')]]$
            \ilyam{add a lemma?};
            \item $[[ [ausr] uM' ]] = [[ [ausr] (uQ' → uM'') ]] = [[ [ausr | uv uQ'] uQ' → [ausr | uv uM''] uM'' ]] =
            [[ [ausr1] uQ' → [ausr2] uM'' ]] = [[ uQ → uM''']] = [[ uM ]]$;
            \item Since $[[ [ausr]β̂⁻]]$ is either equal to  $[[ [ausr1]β̂⁻]]$ or $[[ [ausr2]β̂⁻]]$,
            it inherits their property that it is uniquely determined by $[[ [aus'1]β̂⁻]]$, $[[ [aus'2]β̂⁻]]$, and $[[Γ]]$.
        \end{itemize}

        \item $[[iP1]] = [[iP2]] = [[a⁺]]$. This case is symmetric to \cref{case:anti-unification-initial:nvar}.
        \item $[[iP1]] = [[↓iN1]]$ and $[[iP2]] = [[↓iN2]]$. This case is symmetric to \cref{case:anti-unification-initial:shiftu}
        \item $[[iP1]] = [[∃nas.iP'1]]$ and $[[iP2]] = [[∃nas.iP'2]]$. This case is symmetric to \cref{case:anti-unification-initial:forall}
        \end{caseof}

\end{proof}

>>>>

        \item \ruleref{ottdruleAUArrowLabel}
        In this case, $[[Ξ]] = [[Ξ' ∪ Ξ'']]$, $[[aus1]] = [[aus'1 ∪ aus''1]]$, and 
        $[[aus2]] = [[aus'2 ∪ aus''2]]$,
        where the property holds for ($[[Ξ']]$, $[[aus'1]]$, $[[aus'2]]$) and 
        ($[[Ξ'']]$, $[[aus''1]]$, $[[aus''2]]$) by the induction hypothesis.
        Then since the union of solutions does not change the types the variables are mapped to,
        the required property holds for $[[Ξ]]$, $[[aus1]]$, and $[[aus2]]$.

\begin{lemma}[Soundness of Anti-Unification] \label{lemma:anti-unification-soundness}
    \hfill
    \begin{itemize}
        \item [$+$]  Assuming $[[iP1]]$ and $[[iP2]]$ are normalized,
        if $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$
        then 
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uQ]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$
            are anti-unification solutions, and
            \item $[[ [ausi] uQ = iPi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}

        \item [$-$] Assuming $[[iN1]]$ and $[[iN2]]$ are normalized,
        if $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
        then
        \begin{enumerate}
            \item $[[Γ ; Ξ ⊢ uM]]$,
            \item $[[Γ ; · ⊢ ausi : Ξ]]$ for $i \in \{1,2\}$
            are anti-unification solutions, and
            \item $[[ [ausi] uM = iNi ]]$ for $i \in \{1,2\}$.
        \end{enumerate}
    \end{itemize}
\end{lemma}
\begin{proof}
    We prove it by induction on 
    $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$
    and mutually, $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$.
    Let us consider the last rule applied to infer this judgement.
    \begin{caseof}
        \item \ruleref{\ottdruleAUNVarLabel}, then $[[iN1]] = [[α⁻]] = [[iN2]]$,
              $[[Ξ]] = [[·]]$, $[[uM]] = [[α⁻]]$, and $[[aus1]] = [[aus2]] = [[·]]$.
            \begin{enumerate}
                \item $[[Γ ; · ⊢ α⁻]]$ follows from the assumption $[[Γ ⊢ α⁻]]$,
                \item $[[Γ ; · ⊢ · : ·]]$ holds trivially, and
                \item $[[ [·] α⁻ = α⁻ ]]$ holds trivially.
            \end{enumerate}
        \item \label{case:anti-unification-soundness:shiftu}
         \ruleref{\ottdruleAUShiftULabel}, then $[[iN1]] = [[↑iP1]]$,
                $[[iN2]] = [[↑iP2]]$, and the algorithm makes the recursive call:
                $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
                returning $[[(Ξ, ↑uQ, aus1, aus2)]]$ as the result.

@@

                Notice that the induction hypothesis is applicable to 
                $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$:
                $[[iP1]]$ and $[[iP2]]$ are normalized, since $[[iN1]] = [[iP1 → iN'1]]$
                and $[[iN2]] = [[iP2 → iN'2]]$ are normalized.
                Similarly, the induction hypothesis is applicable to
                $[[Γ ⊨ iN'1 ≈au iN'2 ⫤ (Ξ', uM, aus'1, aus'2)]]$.

@@

                    \item $[[Γ ; · ⊢ ausi : Ξ]]$ and $[[Γ ; · ⊢ aus'i : Ξ']]$
                        Then $[[Γ ; · ⊢ ausi ∪ aus'i : Ξ ∪ Ξ']]$
                        are well-defined anti-unification solution.
                        Let us take an arbitrary $[[β̂⁻]] \in [[Ξ ∪ Ξ']]$.
                        If $[[β̂⁻]] \in [[Ξ]]$.
                        then $[[Γ ; · ⊢ ausi : Ξ]]$ implies that $[[ausi]]$, and hence,
                        $[[ausi ∪ aus'i]]$ contains an entry well-formed in $[[Γ]]$.
                        If $[[β̂⁻]] \in [[Ξ']]$, the reasoning is symmetric.

                        $[[ausi ∪ aus'i]]$ is a well-defined anti-unification solution:
                        any anti-unification variable occurs 
                        uniquely $[[ausi ∪ aus'i]]$, since by \cref{obs:names-defined-by-mapping},
                        the name of the variable is in one-to-one correspondence with 
                        the pair of types it is mapped to
                        by $[[aus1]]$ and $[[aus2]]$, 
                        an is in one-to-one correspondence with the pair of types it is mapped to
                        by $[[aus'1]]$ and $[[aus'2]]$ 
                        i.e.  if $[[β̂⁻]] \in [[Ξ ∩ Ξ']]$ then $[[ [aus1]β̂⁻ = [aus'1]β̂⁻ ]]$,
                        and $[[ [aus2]β̂⁻ = [aus'2]β̂⁻ ]]$.

                    \item $[[ [ausi] uQ = iPi ]]$ and $[[ [aus'i] uM = iN'i ]]$.
                    Since $[[ausi ∪ aus'i]]$ restricted to $[[Ξ]]$ is $[[ausi]]$,
                    and $[[ausi ∪ aus'i]]$ restricted to $[[Ξ']]$ is $[[aus'i]]$,
                    we have $[[ [ausi ∪ aus'i] uQ = iPi ]]$ and 
                    $[[ [ausi ∪ aus'i] uM = iN'i ]]$, and thus, 
                    $[[ [ausi ∪ aus'i] uQ → uM = iP1 → iN'1 ]]$
                \end{enumerate}

@@

@@

\begin{lemma}[Completeness of Anti-Unification] \label{lemma:anti-unification-completeness}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            there exists $[[(Ξ', uQ', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$ 
                are anti-unification solutions, and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

@@

        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            there exists $[[(Ξ', uM', aus'1, aus'2)]]$ such that
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$,
                are anti-unification solutions, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$.
            \end{enumerate}

@@

            Moreover, $[[ [aus'i]uQ' = iPi ]]$ and $[[ [aus'i]uM'' = iN''i ]]$, which means that
            $[[(Ξ', uQ', aus'1, aus'2)]]$ is an anti-unifier of $[[iP1]]$ and $[[iP2]]$,
            and $[[(Ξ', uM'', aus'1, aus'2)]]$ is an anti-unifier of $[[iN''1]]$ and $[[iN''2]]$.
            Then by the induction hypothesis, 
            $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ1, uQ, aus1, aus2)]]$ and 
            $[[Γ ⊨ iN''1 ≈au iN''2 ⫤ (Ξ2, uM, aus3, aus4)]]$ succeed.
            The result of the algorithm is $[[(Ξ1 ∪ Ξ2, uQ → uM, aus1 ∪ aus3, aus2 ∪ aus4)]]$.

@@

\begin{lemma}[Initiality of Anti-Unification] \label{lemma:anti-unification-initial}
    \hfill
    \begin{itemize}
        \item [$+$] 
            Assume that $[[iP1]]$ and $[[iP2]]$ are normalized, and
            $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ, uQ, aus1, aus2)]]$, 
            then $[[(Ξ, uQ, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uQ', aus'1, aus'2)]]$, i.e.
            if 
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uQ']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$
                are anti-unification solutions , and
                \item $[[ [aus'i] uQ' = iPi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uQ')]]$ and $[[ [ausr] uQ' = uQ ]]$. 
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
        \item [$-$] 
            Assume that $[[iN1]]$ and $[[iN2]]$ are normalized, and
            $[[Γ ⊨ iN1 ≈au iN2 ⫤ (Ξ, uM, aus1, aus2)]]$, 
            then $[[(Ξ, uM, aus1, aus2)]]$ is more specific than
            any other sound anti-unifier $[[(Ξ', uM', aus'1, aus'2)]]$, i.e.
            if
            \begin{enumerate}
                \item $[[Γ ; Ξ' ⊢ uM']]$,
                \item $[[Γ ; · ⊢ aus'i : Ξ']]$ for $i \in \{1,2\}$
                are anti-unification solutions, and
                \item $[[ [aus'i] uM' = iNi ]]$ for $i \in \{1,2\}$
            \end{enumerate}
            then there exists $[[ausr]]$ such that
            $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ and $[[ [ausr] uM' = uM ]]$.
            Moreover, $[[ [ausr]β̂⁻]]$ can
            be uniquely determined by $[[ [aus'1]β̂⁻ ]]$, $[[ [aus'2]β̂⁻ ]]$, and
            $[[Γ]]$.
    \end{itemize}
\end{lemma}
\begin{proof}
    First, let us assume that $[[uM']]$ is a metavariable $[[α̂⁻]]$. 
    Then we can take $[[ausr]] = [[α̂⁻]] \mapsto [[uM]]$, which satisfies the required properties:
    \begin{itemize}
        \item $[[Γ ; Ξ ⊢ ausr : (Ξ' | uv uM')]]$ holds since 
        $[[Ξ' | uv uM']] = [[{α̂⁻}]]$ and $[[Γ ; Ξ ⊢ uM]]$ by the soundness of anti-unification (\cref{lemma:anti-unification-soundness});
        \item $[[ [ausr] uM' = uM ]]$ holds by construction
        \item $[[ [ausr]α̂⁻]] = [[uM]]$ is the anti-unifier of 
            $[[iN1]] = [[ [aus'1] α̂⁻]]$ and $[[iN2]] = [[ [aus'2] α̂⁻]]$
            in context $[[Γ]]$, and hence, it is uniquely determined by them (\cref{obs:anti-unification-deterministic}).
    \end{itemize}

@@

        Then by the completeness of anti-unification (\cref{lemma:anti-unification-completeness}),
        the anti-unification algorithm succeeds on $[[iP1]]$ and $[[iP2]]$:
        $[[Γ ⊨ iP1 ≈au iP2 ⫤ (Ξ1, uQ, aus1, aus2)]]$;
        and on $[[iN'1]]$ and $[[iN'2]]$:
        $[[Γ ⊨ iN''1 ≈au iN''2 ⫤ (Ξ2, uM''', aus3, aus4)]]$.
        Notice that $[[aus1 & aus3]]$ and $[[aus2 & aus4]]$ are defined, 
        in other words, for any $[[β̂⁻]] \in [[Ξ1]] \cap [[Ξ2]]$,
        $[[ [aus1] β̂⁻ = [aus2] β̂⁻ ]]$ and $[[ [aus3] β̂⁻ = [aus4] β̂⁻ ]]$,
        which follows immediately from \cref{obs:names-defined-by-mapping}.
        This way, the algorithm proceeds by applying \ruleref{\ottdruleAUArrowLabel} and returns
        $[[(Ξ1 ∪ Ξ2, uQ → uM''', aus1 ∪ aus3, aus2 ∪ aus4)]]$.

@@

        Let us show that $[[ausr]] = [[ausr1 ∪ ausr2]]$ satisfies the required properties:
        \begin{itemize}
            \item $[[Γ ; Ξ1 ∪ Ξ2 ⊢ ausr1 ∪ ausr2 : (Ξ' | uv uM')]]$ holds since 
            $[[Ξ' | uv uM']] = [[Ξ' | uv uQ' → uM'']] = [[(Ξ' | uv uQ') ∪ (Ξ' | uv uM'')]]$,
            $[[Γ ; Ξ1 ⊢ ausr1 : (Ξ' | uv uQ')]]$ and $[[Γ ; Ξ2 ⊢ ausr2 : (Ξ' | uv uM'')]]$;
            \item $[[ [ausr] uM' ]] = [[ [ausr] (uQ' → uM'') ]] = [[ [ausr | uv uQ'] uQ' → [ausr | uv uM''] uM'' ]] =
            [[ [ausr1] uQ' → [ausr2] uM'' ]] = [[ uQ → uM''']] = [[ uM ]]$;
            \item Since $[[ [ausr]β̂⁻]]$ is either equal to  $[[ [ausr1]β̂⁻]]$ or $[[ [ausr2]β̂⁻]]$,
            it inherits their property that it is uniquely determined by $[[ [aus'1]β̂⁻]]$, $[[ [aus'2]β̂⁻]]$, and $[[Γ]]$.
        \end{itemize}

        \item $[[iP1]] = [[iP2]] = [[a⁺]]$. This case is symmetric to \cref{case:anti-unification-initial:nvar}.
        \item $[[iP1]] = [[↓iN1]]$ and $[[iP2]] = [[↓iN2]]$. This case is symmetric to \cref{case:anti-unification-initial:shiftu}
        \item $[[iP1]] = [[∃nas.iP'1]]$ and $[[iP2]] = [[∃nas.iP'2]]$. This case is symmetric to \cref{case:anti-unification-initial:forall}
        \end{caseof}
\end{proof}
